---
title: "Sample Project Tree in R"
author: "mareviv"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    highlight: tango
editor_options: 
  chunk_output_type: console
params:
   dataname: "iris"
   sheet: 1
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE, comment="  ")
```

```{r}
# params <- NULL
# params$dataname <- "iris"
x <- NULL
y <- NULL
fac <- NULL
xlabs <- NULL
ylabs <- NULL
fac_legend <- NULL
```

# Libraries & working directories

```{r results='hide'}
source("./rscripts/libraries.R")
source("./rscripts/directories.R")
source("./rscripts/ggtheme.R")
```

# Data import

```{r}
# source("./rscripts/import_xlsx.R")
# source("./rscripts/import_xls.R")
# source("./rscripts/import_csv.R")
source("./rscripts/import_Rdata.R", echo=TRUE)

# Clean data
# source("./rscripts/clean_data.R") # not necessary in iris dataset
```

# Plotting

```{r}
source("./rscripts/summaries.R", echo=TRUE)
```

### Bar plots 

```{r fig.height=5, fig.width=6}
x <- mydata %>%
  select(Species)
fac <- mydata %>%
  select(Species)

source("./rscripts/barplot.R", echo=TRUE)
barplot1 <- p1
barplot1
```

### Histograms & Density plots

```{r fig.height=5, fig.width=6}
x <- mydata %>%
  select(Sepal.Length)
fac <- mydata %>%
  select(Species)

source("./rscripts/histogram.R", echo=TRUE)
histog1 <- p1 + 
  facet_grid(Species ~ .)
histog1 

source("./rscripts/density.R", echo=TRUE)
densit1 <- themed_basic(p1) + 
  facet_grid(Species ~ .)
densit1 
```

### Scatterplots

```{r fig.height=5, fig.width=6}
x <- mydata %>%
  select(Sepal.Length)
y <- mydata %>%
  select(Sepal.Width)
fac <- mydata %>%
  select(Species)

source("./rscripts/scatterplot.R", echo=TRUE)
scatter1 <- p1
scatter1
```


# Simple hypothesis testing

### Chi-square

Tutorial: [statmethods.net/stats/frequencies.html](https://www.statmethods.net/stats/frequencies.html)

We divide one of our varibles in quantiles: `Petal.Width`  

```{r}
# Define quantiles
quant <- quantile( mydata$Petal.Width, seq(0,1,.25))
# Create new variable from quantiles
mydata$petal_factor <- cut( mydata$Petal.Width, unique(quant),include.lowest=TRUE)
```

```{r}
x <- mydata %>%
  select(Species)
y <- mydata %>%
  select(petal_factor) 
table(x[,1], y[,1]) %>% 
  kable()
chisq_test1 <- table(x[,1], y[,1]) %>%
  chisq.test()
chisq_test1
```


### Wilcoxon rank-sum test

Tutorial: [r-tutor.com/mann-whitney-wilcoxon-test](http://www.r-tutor.com/elementary-statistics/non-parametric-methods/mann-whitney-wilcoxon-test)

```{r}
dat <- mydata %>%
  select(Species, Sepal.Width) %>%
  mutate(Species2 = ifelse(Species == "setosa", "Setosa", "Versicolor/Virginica")) %>%
  mutate(Species2 = as_factor(Species2))

dat %>%
  group_by(Species2) %>%
  summarise(Sepal_width_median = median(Sepal.Width), mean=mean(Sepal.Width), sd=sd(Sepal.Width), count = n()) %>%
  kable()

wilcox <- wilcox.test(dat$Sepal.Width ~ dat$Species2, data=dat)
wilcox
```

### ANOVA & Kruskall Wallis

Turorial: [statmethods.net/stats/anova.html](https://www.statmethods.net/stats/anova.html)

```{r}
dat %>%
  group_by(Species) %>%
  summarise(Sepal_width_median = median(Sepal.Width), mean=mean(Sepal.Width), sd=sd(Sepal.Width), count = n()) %>%
  kable()
```

```{r}
mydata %>%
  select(Species, Sepal.Width, petal_factor) %>%
  group_by(Species, petal_factor) %>%
  summarise(Sepal_width_median = median(Sepal.Width), mean=mean(Sepal.Width), sd=sd(Sepal.Width), count = n()) %>%
  kable()
```

```{r}
krus <- kruskal.test(Sepal.Width ~ Species, data = dat)
krus
```

```{r}
aov1 <- aov(Sepal.Width ~ Species, data = dat)
aov1
summary(aov1) 
aov1$p.value <- summary(aov1)[[1]][[1,"Pr(>F)"]]
```

### Penalisation

```{r comment="  ", message=FALSE, warning=FALSE}
# Construct p-values table
pvalues <- tibble(chi_spec_petal = chisq_test1$p.value, 
                  wil_spec_sepal = wilcox$p.value, 
                  kru_spec_sepal = krus$p.value,
                  aov_spec_sepal = aov1$p.value
                  ) %>%
  t()  
pvalues <-  as.data.frame(pvalues, row.names = rownames(pvalues)) %>%
  mutate(name = rownames(pvalues)) %>%
  rename(p = V1)

source("./rscripts/penalisation.R", echo=FALSE) # uses packages qvalue & kableExtra, gives objects pvalues & pvalues_html

pvalues_html %>%
  kable("html", escape = F, align = "c") %>%
  kable_styling("striped", full_width = F)
```

<div class="alert alert-warning">
Por Benjamini-Hochberg FDR, s&oacute;lo se considerar&aacute;n significativos los p-valores **< `r round(p_value_threshold, 3)`**.   

Por Bonferroni, s&oacute;lo se considerar&aacute;n significativos los p-valores **< `r round(p_value_bonferroni, 5)`**.  

</div>




# Save plots

```{r}
plot_width <- 6
plot_height <- 4

# TO PDF 
plotname_scatter1 <- "scatter1.pdf"
plotname_histog1 <- "histog1.pdf"
plotname_densit1 <- "densit1.pdf"
# TO PNG 
plotname_barplot1 <- "barplot1.png"

routeplot_scatter1 <- file.path(plotdir, plotname_scatter1)
routeplot_histog1 <- file.path(plotdir, plotname_histog1)
routeplot_densit1 <- file.path(plotdir, plotname_densit1)
routeplot_barplot1 <- file.path(plotdir, plotname_barplot1)
ggsave(routeplot_scatter1, scatter1, width = plot_width, height = plot_height) # (see http://ggplot2.tidyverse.org/reference/ggsave.html)
ggsave(routeplot_histog1, histog1, width = plot_width, height = plot_height) 
ggsave(routeplot_densit1, densit1, width = plot_width, height = plot_height)
ggsave(routeplot_barplot1, barplot1, width = plot_width, height = plot_height) 
```


# Save data

```{r}
exported_dataname <- "mydata_exported"
```

```{r}
source("./rscripts/save_data.R", echo=TRUE)
```
